---
title: "PSY3801- Lab 05 Fall 2023"
author: "Amanda Mae Woodward"
date: "2023-10-01"
output: html_document
---
# Welcome back to lab

Last week, we worked on graphing and standard scores. This week, we'll work on understanding probability and its relationship to standard scores. 

## Learning Outcomes:

By the end of this lab, students should be able to: 

- Calculate the probability associated with getting higher or lower than a given standard score
- Calculate the standard score associated with a specific probability 
- Calculate the raw score associated with a specific standard score or probability
- Articulate what a sampling distribution is and explain how the central limit theorem relates 

## Loading the Data: 
the data we are going to use is about loneliness and candy consumption. The data are as follows: 

**participant:** a number between 1 and 50 indicating the participant
**loneliness:** a loneliness score ranging from 1(not at all lonely) to 10 (very lonely)
**candyConsumed:** amount of candy a person ate 
```{r}

```

### Learning Outcome #1: Calculate the probability associated with getting higher or lower than a given standard score

### Reminder on standard score calculation: 
We can use the `scale()` function to calculate standard scores for a whole
variable, if we have access to the whole variable. In our case, we have
the data and can calculate a z score this way.

`scale()` uses the following format: `scale(data$columnName)`

Let's calculate the standard score for the loneliness scale
```{r}
candyDat$lonelinessZ<- scale(candyDat$loneliness)
```

We can view these numbers by looking at our dataset and compare them to our mean and standard deviation.  

Take a minute to talk to each other about interpretation: 
1) what does a z score of 0.29 mean? 
2) what does a z score of -1.17 mean? 
3) what does a z score of 1.39 mean? 

In addition to saying which standard score is higher or lower, we can look at probability. Probability allows us to ask questions like: 
 
 "What is the likelihood of getting a standard score of -1?" 
 
To answer this questions, we need to use the function `pnorm()`

`pnorm()` gives the probability of getting a range of scores. We can decide if we want higher or lower than a specific score. 

Let's say that we want to know the probability of obtaining a score less than or equal to 1.75. We would use the following: 

`pnorm(zScore, lower.tail = TRUE/FALSE)`

```{r}
library(ggplot2)
ggplot(, aes(rnorm(100000)))+ geom_histogram()+ geom_vline(xintercept = 1.75,color="purple", size=1)+theme_classic()
```

```{r}
pnorm(1.75, lower.tail= TRUE)

```

We can look at the graph we made above and see the area represented by this probability. 

If we wanted to get the probability of 1.75 or higher, we could do this
```{r}
pnorm(1.75, lower.tail=FALSE)
```

Remember, the probability of the whole distribution is equal to 1. If you wanted to, you could use the answer from "less than" and subtract it from 1 to get the same answer.
```{r}
1- 0.9599408
```

### Learning Outcome 1 Practice
1) calculate the probability of obtaining a z score of 0 or less
2) calculate the probability of obtaining a z score below -1.5 
3) calculate the probability of obtaining a z score above 1.5

### Learning Outcome #2: Calculate the standard score associated with a specific probability

We can also work in the opposite direction. We can use a probability of a score to look up a specific z score or standard score using the function `qnorm()` In other words, we're asking what standard score is the boundary or cut off for a specific area, or probability 

Let's find the standard score (z score) associated with a probability of .09
`qnorm(probability, lower.tail = TRUE/FALSE)`
```{r}
qnorm(.09, lower.tail = TRUE )
```
This gives us the score associated with a probability of .09 or less. 

If we want to know about the highest .09, we would need to do the following:
```{r}
qnorm(.09, lower.tail=FALSE)
```

### Learning Outcome 2 Practice: 
1) What z score cuts off the top 5% (0.05) of the distribution? 
2) What z score is associated with probability of .46 in in the lower side?

### Learning Outcome #3: Calculate the raw score associated with a specific standard score or probability

The last thing that we need to talk about with standard scores is how we can go from a standard score to the value in raw score units. 

z= score - mean/ standard deviation 

z * standard deviation + mean = raw score

Let's practice with the Valentine's day data. 

We want to know how much candy the top 20% of consumers eat on the holiday. To look at this, we need to first find the z score associated with the highest 20%
```{r}
z<- qnorm(.2, lower.tail=FALSE)
z
```

Then, we need to transform our standard score of .84 to the raw score. To do this, we need to use another version of the equation: 
z * standard deviation + mean = raw score
```{r}
z* sd(candyDat$candyConsumed) + mean(candyDat$candyConsumed)
```
The top 20% eat at least 20 pieces of candy (rounded up). 

### Learning Outcome #3 Practice: 
1) How much candy do the lowest 10% of participants eat? 
2) How much candy do the middle 50% eat? 


