---
title: "PSY3801- Lab Week 3 Descriptive Statistics"
author: "Amanda Mae Woodward"
date: "2023-09-16"
output: html_document
---
# Welcome back to lab! 

This week, we will go over how to calculate descriptive statistics. This will allow us to build off of our R foundations and talk about our data meaningfully. 

## Learning Outcomes:

By the end of this lab, students should be able to:
1) calculate and interpret measures of central tendency
2) calculate and interpret measures of variability


Before we go through descriptive statistics, we need to upload some data. We're going to use data about the amount of candy consumed around Halloween. We have three variables: 

participant: the participant number of each person responding
candyType: the candy that is being rated
candyRating: the rating of the candy (1= not like at all, 5= really like)
candyConsumed: the number of pieces of candy consumed

Try to import the data on your own, then we will go over it together. 
```{r}
#copy and paste the dataset name<-read.csv() code here
```


### Learning Outcome #1: calculate and interpret measures of central tendency

#### Mean
To calculate the mean, we use the function `mean()`. We'll calculate the mean of "candyConsumed" 

`mean(dataset$column)`
```{r}


```

#### Median
To calculate the median, we use the function `median()` We'll calculate the median of  "candyConsumed" 

`median(data$columnname)`

```{r}


```

#### Mode
This one is a little different. The command "mode" means something else in R, so we need to use the package we installed last week,"modeest." Then we can use the command `mfv()` to find the mode.

 We'll calculate the mode of "candyConsumed"
 
As an aside, `mode(data$columnname)`will not tell us what value occurs the most. It will tell us what type of data are stored in R. 

`mfv(data$columnname)`

```{r}
library(modeest)

```

#### Central Tendency Practice
In order to practice the measures of central tendency, we'll use the candy rating variable. While it is technically ordinal, we will treat it as interval for this example. 

1) calculate the mean of the candyRating variable. What does this value mean?
2) calculate the median of the candyRating variable. What does this value mean?
3) calculate the mode of the candyRating variable. What does this value mean? 

### Learning Outcome #2: Calculate and interpret measures of variability

#### Range
To calculate the range, use `range()`.  We'll calculate the range of "candyConsumed"
`range(data$columnname)`
```{r}

```
This gives you the highest and lowest values. Then, you need to subtract the lowest from the highest value. 

Our range is: 

```{r}

```

#### Interquartile Range
To calculate the interquartile range, we use `IQR()`.  We'll calculate the IQR of "candyConsumed"

`IQR(data$columnname)`

```{r}

```

#### Standard deviation

To calculate the standard deviation, we use `sd()`.  We'll calculate the standard deviation of "candyConsumed"

`sd(data$columnname)`

```{r}

```

#### Variance

To calculate the variance, we use `var()` or `sd()^2`  We'll calculate the variance of "candyConsumed"

`var(data$columnname)`
```{r}

```

#### Variability Practice
1) calculate the range of the candyRating variable. What does this value mean? 
2) calculate the IQRof the candyRating variable. What does this value mean? 
3) calculate the standard deviation of the candyRating variable. What does this value mean?  
4) calculate the variance of the candyRating variable. What does this value mean? 

